---
title: "HindFootLength"
format: html
editor: visual
code-fold: true 
---

## Hind-foot Length

## Loading Packages

```{r}
library(tidyverse)
library(magrittr)
```

# Loading External Survey Data

```{r}
download.file(url = "https://ndownloader.figshare.com/files/2292169",
              destfile = "data/Survey_data.csv")
survey <- read.csv("data/Survey_data.csv")
```

# Calculating Mean Hind-foot Length

```{r}
#| echo: false
mean(survey$hindfoot_length,na.rm=TRUE)
```

# Manipulating Survey Data to Calculate Mean Yearly Hind-foot Length

```{r}

mean_hindfoot <-
  survey %>%
  filter(!is.na(hindfoot_length)) %>%
  mutate(hindfoot_length_cm = hindfoot_length /10) %>%
  group_by(year) %>%
  summarize(mean_hindfoot_cm = mean(hindfoot_length_cm))
```

# Mean Yearly Hind-foot Length

```{r}
#| echo: false

plot(mean_hindfoot)
```

# Species-specific weights through time

```{r}
#| echo: false  
rodents <- survey %>% dplyr::filter(taxa == "Rodent") %>%
  group_by(species_id, year) %>% 
  dplyr::filter(!is.na(weight)) %>%
  summarise(mass = mean(weight),
            n = n(),
            se_w = sd(weight)/sqrt(n))


```

# Mean yearly weight by species

```{r}
#| echo: false

ggplot(rodents, aes(x = year, y = mass, ymin = mass - se_w, ymax = mass + se_w, group = species_id)) +
  geom_point() +
  geom_errorbar() +
  geom_line() +
  facet_wrap(. ~ species_id) +
  theme_bw()

ggsave(filename = "graphics/faceted_mass.png",
	width = 20,
	height = 20,
	dpi = 320
	)
```

Several of these species don't really have enough obs through the time series to merit including them. Let's remove some

```{r}
#| echo: false

sums <- rodents %>% group_by(species_id) %>%
  summarise(n_tot = sum(n))

```

Of the 25 rodent species in order, there's a jump from n = 141 to n = 876, feels like a natural breakpoint between "common" and not. And leaves us 12 species with a lot of values, let's focus on them and see what the plots look like in that case

```{r}
#| echo: false

incl <- c(subset(sums, n_tot > 800)$species_id)

rodents %<>% filter(species_id %in% incl)

ggplot(rodents, aes(x = year, y = mass, ymin = mass - se_w, ymax = mass + se_w, colour = species_id)) +
  geom_point() +
  geom_errorbar() +
  geom_line() +
  ylab("Mass (g)") +
  viridis::scale_colour_viridis(discrete = TRUE, option = "C") +
 # facet_wrap(. ~ species_id) +
  theme_bw()
```

NL = Neotoma albigula is really variable! the rest of them seem fairly flat.
